### 2016-02-02 docs/response_to_referee.txt
Added repsonse to main comment (and a few minor ones)

We would like to thank the referee for their very useful comments.
We have included response to them inline below and make note any
changes to the manuscript. These changes are displayed in red in the
resubmitted version of the manuscript.

Major comment: "In section 5, I have a problem with the assumption of the sGRB jet
opening angle being modelled as a half-normal distribution. Figure 3
in Fong et al. (2014) shows that the long GRB distribution may be
modelled by something like a Rayleigh distribution (or something
similar to give a tail to high values), but with little weight at
small angles. If the same is true of the sGRB population (and from
Figure 3 the long tail seems to be present here too) then a
half-normal distribution for the opening angle is not a good match to
the data. Did you try other priors on the inclination angle which may
better represent the true distribution? I ask that you try to use a
better prior for the inclination angle, or alternatively provide me
with a better justification of your current choice."

Response: Our poor description regarding how we use the opening angle (theta_j)
distribution to constrain the distribution on the orientation angle (iota) of
the systems has caused confusion here. We are not really interested
in the jet opening angle distribution, but are just trying to somehow use
it as a proxy for getting an idea of the orientation angle distribution,
which would not be expected to look the same. Under the assumption that to
observe a sGRB we must have theta_j >= |iota|, and assuming a Rayleigh
distribution on theta_j, the distribution on |iota| does become a half-normal
distribution. In fact our choice of a half-normal for |iota| with a sigma value of
14.8 degs is rather conservative (in that it extends to larger iota values
[this is something we would plan to correct in future studies]) compared to the
half-normal distribution that you truly get based a Rayleigh distribution in 
theta_j with a median of 10 degs. We have added text to the manuscipt to
hopefully make this clearer. If the referee would like to see more detail on
these we can send them further information.

Minor comments
--------------

Comment: "The first sentence should be followed by a reference to Abadie et
al. (2010), or a similar updated article on predicted detection rates
if available."

Response:

Comment: ""will be detected out to a maximum range of ≈ 450 Mpc." -> clarify
whether this is the horizon distance or sensemon distance."

Response:

Comment: ""Observations of sGRBs give an inferred rate of sGRB producing CBC
mergers" -> mention that the sGRB rate must be corrected for beaming
effects to map to CBC merger rate."

Response:

Comment: ""may be possible to identify the host galaxy of the source" -> is a
host galaxy necessary for redshift identification? Can the redshift
not be attained from spectroscopic analysis of the sGRB light curve?"

Response:

Comment: ""With knowledge of the redshift, using current best estimates for
the Hubble constant and other cosmological parameters, the true
luminosity distance can be estimated to ∼ 1% accuracy." -> reference
needed for current estimates of cosmological parameters."

Response:

Comment: "Semi-colon in first sentence can be simply replaced by a comma."

Response:

Comment: ""at the frequency range used in this project" --> explicitly
mention the frequency range."

Response:

Comment: ""error in the amplitude calibration" -> is this the strain
amplitude or the error signal output. Clarify either way."

Response:

Comment: ""within a control feedback loop (see [7]), which are subject to
uncertainties." -> This is quite a vague sentence. What kind of
uncertainties?"

Response:

Comment: "Even though the math is quite simple, I would advocate explicitly
stating the relationship between the true length response function and
the estimated one via the calibration scale factor."

Response:

Comment: ""larger amplitude (e.g. be closer)" -> it would only appear to be
closer if we had a handle on the other parameters affecting the
amplitude e.g. the chirp mass. We obviously get the latter from the
signal's phase evolution, but this point should be clarified."

Response:

Comment: "How realistic is the assumption of a constant calibration scale
factor over the full LIGO band? What physical mechanism would lead us
to make a poor estimation of the length response function, and why
would other calibration tests fail to mitigate the effect dealt with
in this article?"

Response:

Comment: ""different signal arrival times at each detector in the network" ->
the concept of a network has not been mentioned or discussed before. A
simple sentence explaining the purpose and composition of the kind of
detector network expected in the advanced era is necessary."

Response:

Comment: "A reference for the Einstein Telescope is necessary, whether this
be the original design study, or possibly studies of arm orientation
preferences."

Response:

Comment: ""The use of GW events with sGRB counterparts for use as a
cosmological tool was investigated in [11] and [12] with respect to
the 3rd generation GW interferometer, the Einstein Telescope (ET)." ->
the concept was also explored for advanced era detectors in a later
section of Taylor, Gair, Mandel (2012)."

Response:

Comment: ""These instead use either the distribution [...] assumed neutron
star (NS) mass" -> to clarify, Taylor, Gair, Mandel (2012) assume the
*form* of the NS mass distribution, but search over the parameters of
the distribution as hyperparameters in a hierarchical modeling scheme."

Response:

Comment: ""specifically for the Hubble constant, which is has uncertainties
of O(1%)" -> reference to Planck results or other appropriate sources."

Response:

Comment: ""If the event exhibits an optical afterglow then the host galaxy
can be identified and from which a redshift can be obtained" -> I
reiterate my question from earlier: why can the redshift not be
determined from the sGRB light curve itself?"

Response:

Comment: ""The fraction of events with associated redshifts is ∼1/3" -> do
you have a reference for this?"

Response:

Comment: ""equal to the horizon range of the advanced GW detector network for
BNS systems." -> what is the network composition in this case?"

Response:

Comment: ""derived from three main sources, population synthesis models..."
-> "derived from three main sources: population synthesis models...""

Response:

Comment: ""In [3] using the rate of observed sGRBs and assuming a beaming
angle of 15◦ a similar event rate of 1–180 is obtained." -> This
sentence was a bit difficult to parse. I suggest restructuring it to
with something like "A similar event rate of 1-180 was obtained in
[3], where the rate of observed sGRBs was converted to the rate of BNS
events by assuming a beaming angle of 15 degrees.""

Response:

Comment: "Is the beaming angle here the full opening angle, or the half
opening angle?"

Response:

Comment: ""For the GW followup of sGRB scenario the merger time for BNS/NSBH
systems will be estimated from the sGRB to within a few seconds" ->
This sentence was difficult to understand. Suggest changing it to "For
the scenario of a GW-followup of an observed sGRB, the merger time for
BNS/NSBH systems will be estimated from the sGRB to within a few
seconds.""

Response:

Comment: ""However, as discussed earlier, the property of beaming severely
impacts the probable rate of such joint observations." -> So am I to
understand from this final paragraph that although the probability of
such joint observations is low, the SNR distribution of any detected
signals will be comparatively higher than the regular
GW-detection-only SNR distribution?"

Response:

Comment: "In equation 3, I believe the inclination term in hx should simply
be cosine(iota) rather than the square. Please proof check all
equations to ensure they match your intent."

Response: The referee is correct. This has been fixed.

Comment: ""the sky position of the source α and δ." -> I assume these
correspond to RA and DEC, but this should be made clear."

Response: alpha and delta have now been defined as right ascension and declinarion respectively.

Comment: ""for the black hole we use a prior based on the canonical mass
distribution used for the rate results in [34] with a mean of 5 M and
a standard deviation of 1 M." -> you mention that this distribution is
only an example for this type of system, but how would you do this in
reality? Would you make the mean and standard deviation further
parameters in your search in a hierarchical modelling scheme? How
would your current choices bias calibration attempts if the population
you are targeting is significantly different than your assumptions
here?"

Response:

Comment: "In equation 7, do your choices mean p(C=0.1) = p(C=10) or
p(lnC=ln(0.1)) = p(lnC=ln(10))?"

Response: The choice mean p(C=0.1) = p(C=10). This has been fixed.

Comment: ""uniform distribution on the sky" -> does this mean uniform in
right ascension and uniform in cosine(90-declination)?"

Response:

Comment: "Your range of injected calibration scale factors is quite narrow,
meaning you only really tested situations where the factor was pretty
close to optimal. Did you test cases where the detectors were very
poorly calibrated? If not, then I highly suggest that you do so, or
adequately comment on expectations from these tests."

Response:

Comment: ""calculated by averaging 64 separate noisy realisations of the PSD
and scaled with the same calibraion scale factor as applied to the
injection and noise" -> firstly the typo in "calibration" should be
corrected. Also, does this mean that the calibration scale factor
appears in "S" as a parameter also during your MCMC? If so, why does
it not explicitly appear in the denominator of the exp[] of equation
6?"

Response: The typo has been corrected. The calibration scale factor does
not appear in "S" during the MCMC or the denominator of equation 6. The
description in this section is saying how the simulated data was created,
whereas during the MCMC it is the only the expect signal waveform
that has the scale factor applied to it to try and match the data (to which
the calibration factor has already been applied during its creation).

Comment: "In the paragraph on page 6 beginning "In figure 3..." there is some
explanation of why the H1 and L1 uncertainties saturate out to ~25%.
The explanation did not seem very clear to me, and does not address
why V1 does not show a similar saturation. What is the significance of
the value of ~25%? Is there some influence of the finiteness of the
detected population at larger distances that is influencing the width
of the uncertainty distribution?"

Response:

Comment: "In figure 1 the tick labels for the chirp mass are overlaid with
the label. Better labelling or figure formatting is needed here, as
well as units where appropriate. I like the different edge coloring
for the relevant scale factor boxes, but perhaps these red edges could
be thickened for black and white printing."

Response:

Comment: "No mention is given to the number of MCMC iterations, the length of
burn-in, or how convergence was ascertained. A brief comment would be
appropriate."

Response:

Comment: "Page 9, second paragraph. "this prior is not that same" -> "this
prior is not the same""

Response: This has been fixed.